Retrofitting Effect Handlers onto OCaml

### 2020-11-17 retro-concurrency.tex
edits

Kawahara et al.~\cite{Kawahara20} implement one-shot effect handlers using
coroutines as a macro-expressible translation, and present an embedding in Lua
and Ruby. Lua provides asymmetric coroutines~\cite{Lua} where each coroutine
uses its own stack similar to how each handled computation runs in its own
fiber in Multicore OCaml. Lua does not allow coroutines to include C frames, as in
Multicore OCaml. It is useful to note that Lua is an interpreted
language, unlike OCaml, and DWARF unwind tables are not a concern in Lua.

@@

The interaction of non-local control flow and resources has been studied
extensively. Scheme uses |dynamic-wind|~\cite{R5RS}, which is a generalisation of
Common Lisp |unwind-protect|~\cite{Steele90}, which ensures de-allocation (and
re-allocation) of resources every time the non-local control leaves and enters
back into a context. |dynamic-wind| is not quite what one wants for resource
clean up~\cite{Kiselyov,Sitaram03}. The main challenge is distinguishing
returning exits from non-returning ones. Resource clean up needs to occur only
during non-returning exit.

>>>>

Kawahara et al.~\cite{Kawahara20} implement one-shot effect handlers using
coroutines as a macro-expressible translation, and present an embedding in Lua
and Ruby. Lua provides asymmetric coroutines~\cite{Lua} where each coroutine
uses its own stack similar to how each handled computation runs in its own
fiber in Multicore OCaml. Lua does not allow coroutines to include C frames, as in
Multicore OCaml. 
%It is useful to note that Lua is an interpreted
%language, unlike OCaml, and DWARF unwind tables are not a concern in Lua.

@@

The interaction of non-local control flow and resources has been studied
extensively. Scheme uses |dynamic-wind|~\cite{R5RS}, which is a generalisation of
Common Lisp |unwind-protect|~\cite{Steele90}, which ensures de-allocation (and
re-allocation) of resources every time the non-local control leaves and enters
back into a context. |dynamic-wind| is not quite sufficient for resource
clean up~\cite{Kiselyov,Sitaram03} as returning exits must be distinguished
from non-returning ones, and resource cleanup is only necessary during
non-returning exits.
